---
title: "ksfloor_analysis"
author: "BiancaRius"
date: "2026-05-21"
output: html_document
---

```{r}
knitr::opts_chunk$set(echo = TRUE)
# Loading required packages
library(tidyverse)
library(data.table) 
library(fs)         
```

```{r}
# Load data
# script to generate the .rds files: "~/Desktop/Postdoc_Toulouse/Postdoc_Toulouse/TROLL/TROLL_code_understanding_documenting/runs/ksfloor/save_data_rds.R"

# You can change here if you need to assess other analysis
# base_dir <- "/Users/biancarius/Desktop/Postdoc_Toulouse/Postdoc_Toulouse/TROLL/TROLL_code_understanding_documenting/runs/ksfloor/veg/deep_WT/sandy"

base_dir <- "/Users/biancarius/Desktop/Postdoc_Toulouse/Postdoc_Toulouse/TROLL/TROLL_code_understanding_documenting/runs/ksfloor/veg/deep_WT/sandy"

cache_dir <- path(base_dir, "_rds")

plot_dir <- path(base_dir, "plots")
# dir_create(plot_dir)

# Daily data
# df_water_balance_daily <- readRDS(path(cache_dir, "df_water_balance_daily.rds"))
# df_vertical_flux_daily <- readRDS(path(cache_dir, "df_vertical_flux_daily.rds"))

# Annual data
df_water_balance_annual <- readRDS(path(cache_dir, "df_water_balance_annual.rds"))
df_vertical_flux_annual <- readRDS(path(cache_dir, "df_vertical_flux_annual.rds"))
df_biogem_annual <- readRDS(path(cache_dir, "df_biogeochemical_annual.rds"))
```

Full Timeline and Stabilization Point

Before looking at the final state, we need to determine *if* and *when* the system stabilizes. A hydrological model is considered stable (having surpassed the spin-up phase) 

# The Full 274-Year Trajectory

First, let's plot the entire time series to visually identify the spin-up period and whether the curves flatten out.

To consider in the analysis:
Increasing ksfloor reduces the risk of numerical hydraulic locking by preventing extremely low hydraulic conductivity values. However, higher ksfloor values also impose an increasingly artificial lower bound on soil-water movement. If the imposed floor is too high, vertical connectivity may be maintained even when the soil profile should physically become hydraulically disconnected, especially under deep water table and sandy soil conditions. Therefore, the sensitivity analysis should identify the lowest ksfloor that stabilizes the model without artificially increasing vertical water exchange.

```{r plot-full-swc-trajectory}
swc_full_allks <- df_water_balance_annual %>%
  filter(variable == "SWC") %>%
  ggplot(aes(x = year, y = mean_value, color = ksfloor_factor)) +
  geom_line(linewidth = 0.7) +
  facet_wrap(~ layer, scales = "free_y") +
  labs(
    title = "SWC",
    x = "Simulation year",
    y = "Annual mean SWC",
    color = "ksfloor"
  ) +
  theme_bw()

swc_full_allks

# ggsave(
  # filename = path(plot_dir, "SWC_fullTrajectory_allKsfloor.png"),
  # plot = swc_full_allks,
  # width = 10,
  # height = 7,
  # dpi = 300
# )
```

```{r plot-full-swp-trajectory}

swp_full_allks <-  df_water_balance_annual %>%
  filter(variable == "SWP") %>%
  ggplot(aes(x = year, y = mean_value, color = ksfloor_factor)) +
  geom_line(linewidth = 0.7) +
  facet_wrap(~ layer, scales = "free_y") +
  labs(
    title = "SWP",
    x = "Simulation year",
    y = "Annual mean SWP",
    color = "ksfloor"
  ) +
  theme_bw()

swp_full_allks

# ggsave(
#   filename = path(plot_dir, "SWP_fullTrajectory_allKsfloor.png"),
#   plot = swp_full_allks,
#   width = 10,
#   height = 7,
#   dpi = 300
# )

```

```{r plot-full-transpiration-trajectory}

transpiration_full_allks <- df_water_balance_annual %>%
  filter(variable == "transpiration") %>%
  ggplot(aes(x = year, y = mean_value, color = ksfloor_factor)) +
  geom_line(linewidth = 0.7) +
  facet_wrap(~ layer, scales = "free_y") +
  labs(
    title = "Transpiration",
    x = "Simulation year",
    y = "Annual mean transpiration",
    color = "ksfloor"
  ) +
  theme_bw()

transpiration_full_allks

# ggsave(
#   filename = path(plot_dir, "transpiration_fullTrajectory_allKsfloor.png"),
#   plot = transpiration_full_allks,
#   width = 10,
#   height = 7,
#   dpi = 300
# )

```


As 1e_10 seems very extreme, I will plot the other ones

```{r}
ksfloor_to_plot <- c("1e_12", "1e_14", "1e_16", "1e_18", "1e_20")

df_ksfloor_to_plot <- df_water_balance_annual %>% filter(ksfloor_factor %in% ksfloor_to_plot)
```

```{r}
swc_full_ksgt10 <- df_ksfloor_to_plot %>%
  filter(variable == "SWC") %>%
  ggplot(aes(x = year, y = mean_value, color = ksfloor_factor)) +
  geom_line(linewidth = 0.7) +
  facet_wrap(~ layer, scales = "free_y") +
  labs(
    title = "SWC",
    x = "Simulation year",
    y = "Annual mean SWC",
    color = "ksfloor"
  ) +
  theme_bw()

swc_full_ksgt10

# ggsave(
#   filename = path(plot_dir, "swc_fullTrajectory_Ksfloor_gt10.png"),
#   plot = swc_full_ksgt10,
#   width = 10,
#   height = 7,
#   dpi = 300
# )

```

```{r}
swp_full_ksgt10 <- df_ksfloor_to_plot %>%
  filter(variable == "SWP") %>%
  ggplot(aes(x = year, y = mean_value, color = ksfloor_factor)) +
  geom_line(linewidth = 0.7) +
  facet_wrap(~ layer, scales = "free_y") +
  labs(
    title = "SWP",
    x = "Simulation year",
    y = "Annual mean SWP",
    color = "ksfloor"
  ) +
  theme_bw()

swp_full_ksgt10

# ggsave(
#   filename = path(plot_dir, "swp_fullTrajectory_Ksfloor_gt10.png"),
#   plot = swp_full_ksgt10,
#   width = 10,
#   height = 7,
#   dpi = 300
# )

```

```{r}

transpiration_full_ksgt10 <- df_ksfloor_to_plot %>%
  filter(variable == "transpiration") %>%
  ggplot(aes(x = year, y = mean_value, color = ksfloor_factor)) +
  geom_line(linewidth = 0.7) +
  facet_wrap(~ layer, scales = "free_y") +
  labs(
    title = "Transpiration",
    x = "Simulation year",
    y = "Annual mean transpiration",
    color = "ksfloor"
  ) +
  theme_bw()

transpiration_full_ksgt10

# ggsave(
#   filename = path(plot_dir, "transpiration_fullTrajectory_Ksfloor_gt10.png"),
#   plot = transpiration_full_ksgt10,
#   width = 10,
#   height = 7,
#   dpi = 300
# )


```

Plotting layer interface variables

```{r}
mean_ksgeometric_full_allks <- df_vertical_flux_annual %>%
  filter(metric == "mean_ks_harmonic") %>% # actually this is the ks geometric
  ggplot(aes(x = year, y = mean_value, color = ksfloor_factor)) +
  geom_line(linewidth = 0.7) +
  facet_wrap(~ interface, scales = "free_y") +
  scale_y_log10() +
  labs(
    title = "Mean harmonic hydraulic conductivity by interface",
    subtitle = "Annual mean values over the full 274-year simulation",
    x = "Simulation year",
    y = "Annual mean Ks geometric",
    color = "ksfloor"
  ) +
  theme_bw()

mean_ksgeometric_full_allks

ggsave(
  filename = path(plot_dir, "mean_ksgeometric_fullTrajectory_allKsfloor.png"),
  plot = mean_ksgeometric_full_allks,
  width = 10,
  height = 7,
  dpi = 300
)

```


```{r}
gross_volumetricchange_full_allks <- df_vertical_flux_annual %>%
  filter(metric == "gross_volumetric_change") %>%
  ggplot(aes(x = year, y = sum_value, color = ksfloor_factor)) +
  geom_line(linewidth = 0.7) +
  facet_wrap(~ interface, scales = "free_y") +
  labs(
    title = "Gross vertical water exchange by interface",
    subtitle = "Annual sums over the full 274-year simulation",
    x = "Simulation year",
    y = "Annual gross volumetric change",
    color = "ksfloor"
  ) +
  theme_bw()

gross_volumetricchange_full_allks

ggsave(
  filename = path(plot_dir, "gross_volumetricchange_fullTrajectory_allKsfloor.png"),
  plot = gross_volumetricchange_full_allks,
  width = 10,
  height = 7,
  dpi = 300
)

```

As 1e_10 seems very extreme, I will plot the other ones

```{r}
gross_volumetricchange_full_ksgt10 <- df_vertical_flux_annual %>%
  filter(
    metric == "gross_volumetric_change",
    ksfloor_factor != "1e_10"
  ) %>%
  ggplot(aes(x = year, y = sum_value, color = ksfloor_factor)) +
  geom_line(linewidth = 0.7) +
  facet_wrap(~ interface, scales = "free_y") +
  labs(
    title = "Gross vertical water exchange by interface",
    subtitle = "Annual sums excluding ksfloor = 1e-10",
    x = "Simulation year",
    y = "Annual gross volumetric change",
    color = "ksfloor"
  ) +
  theme_bw()

gross_volumetricchange_full_ksgt10

# ggsave(
#   filename = path(plot_dir, "gross_volumetricchange_fullTrajectory_Ksfloor_gt10.png"),
#   plot = gross_volumetricchange_full_ksgt10,
#   width = 10,
#   height = 7,
#   dpi = 300
# )

```

```{r}
# Log scale
gross_volumetricchange_full_ksgt10_log <- df_vertical_flux_annual %>%
  filter(metric == "gross_volumetric_change",
         ksfloor_factor != "1e_10") %>%
  mutate(
    sum_value_plot = if_else(sum_value <= 0, NA_real_, sum_value)
  ) %>%
  ggplot(aes(x = year, y = sum_value_plot, color = ksfloor_factor)) +
  geom_line(linewidth = 0.7, na.rm = TRUE) +
  facet_wrap(~ interface, scales = "free_y") +
  scale_y_log10() +
  labs(
    title = "Gross vertical water exchange by interface",
    subtitle = "Annual sums, log scale",
    x = "Simulation year",
    y = "Annual gross volumetric change, log scale",
    color = "ksfloor"
  ) +
  theme_bw()

gross_volumetricchange_full_ksgt10_log

# ggsave(
#   filename = path(plot_dir, "gross_volumetricchange_fullTrajectory_Ksfloor_gt10_log.png"),
#   plot = gross_volumetricchange_full_ksgt10_log,
#   width = 10,
#   height = 7,
#   dpi = 300
# )

```


```{r}
net_volumetricchange_full_ksgt10 <- df_vertical_flux_annual %>%
  filter(
    metric == "net_volumetric_change",
    ksfloor_factor != "1e_10"
  ) %>%
  ggplot(aes(x = year, y = sum_value, color = ksfloor_factor)) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_line(linewidth = 0.7) +
  facet_wrap(~ interface, scales = "free_y") +
  labs(
    title = "Net vertical water exchange by interface",
    subtitle = "Annual sums excluding ksfloor = 1e-10",
    x = "Simulation year",
    y = "Annual net volumetric change",
    color = "ksfloor"
  ) +
  theme_bw()

net_volumetricchange_full_ksgt10

# ggsave(
#   filename = path(plot_dir, "net_volumetricchange_fullTrajectory_Ksfloor_gt10.png"),
#   plot = net_volumetricchange_full_ksgt10,
#   width = 10,
#   height = 7,
#   dpi = 300
# )


```

Looking to directionality
```{r}
df_directionality_annual <- df_vertical_flux_annual %>%
  filter(metric %in% c("net_volumetric_change", "gross_volumetric_change")) %>%
  select(
    scenario,
    ksfloor_val,
    ksfloor_factor,
    year,
    interface,
    layer_upper,
    layer_lower,
    metric,
    sum_value
  ) %>%
  pivot_wider(
    names_from = metric,
    values_from = sum_value
  ) %>%
  mutate(
    abs_net_volumetric_change = abs(net_volumetric_change),
    directionality = abs_net_volumetric_change / gross_volumetric_change,
    directionality = if_else(
      is.finite(directionality),
      directionality,
      NA_real_
    )
  )

volume_directionality_full_ksgt10 <- df_directionality_annual %>%
  filter(ksfloor_factor != "1e_10") %>%
  ggplot(aes(x = year, y = directionality, color = ksfloor_factor)) +
  geom_line(linewidth = 0.7) +
  facet_wrap(~ interface, scales = "free_y") +
  labs(
    title = "Directionality of vertical water exchange by interface",
    subtitle = "|net volumetric change| / gross volumetric change, excluding ksfloor = 1e-10",
    x = "Simulation year",
    y = "Directionality",
    color = "ksfloor"
  ) +
  theme_bw()

volume_directionality_full_ksgt10

# ggsave(
#   filename = path(plot_dir, "volume_directionality_fullTrajectory_Ksfloor_gt10.png"),
#   plot = volume_directionality_full_ksgt10,
#   width = 10,
#   height = 7,
#   dpi = 300
# )

# directionality ≈ 1
# fluxo anual quase totalmente em uma direção.
# 
# directionality ≈ 0
# há troca nos dois sentidos que se cancela, ou net muito pequeno em relação ao gross.
# 
# directionality = NA
# gross volumetric change é zero ou quase zero, então não há troca suficiente para calcular.

```


```{r combined-interface-annual-table}
# Combine the main annual interface diagnostics in one table

df_interface_diagnostics_annual <- df_vertical_flux_annual %>%
  filter(
    metric %in% c(
      "mean_ks_harmonic",
      "gross_volumetric_change",
      "net_volumetric_change"
    )
  ) %>%
  mutate(
    metric_value = case_when(
      metric == "mean_ks_harmonic" ~ mean_value,
      metric %in% c("gross_volumetric_change", "net_volumetric_change") ~ sum_value,
      TRUE ~ NA_real_
    )
  ) %>%
  select(
    scenario,
    ksfloor_val,
    ksfloor_factor,
    year,
    interface,
    layer_upper,
    layer_lower,
    metric,
    metric_value
  ) %>%
  pivot_wider(
    names_from = metric,
    values_from = metric_value
  ) %>%
  left_join(
    df_directionality_annual %>%
      select(
        scenario,
        ksfloor_val,
        ksfloor_factor,
        year,
        interface,
        directionality
      ),
    by = c("scenario", "ksfloor_val", "ksfloor_factor", "year", "interface")
  )

glimpse(df_interface_diagnostics_annual)

df_interface_diagnostics_annual %>%
  count(ksfloor_factor, interface)
```

```{r summary-interface-diagnostics-full-run}
interface_diagnostics_summary_full <- df_interface_diagnostics_annual %>%
  group_by(
    scenario,
    ksfloor_factor,
    ksfloor_val,
    interface,
    layer_upper,
    layer_lower
  ) %>%
  summarise(
    n_years = n(),
    
    median_ks_harmonic = median(mean_ks_harmonic, na.rm = TRUE),
    min_ks_harmonic = min(mean_ks_harmonic, na.rm = TRUE),
    max_ks_harmonic = max(mean_ks_harmonic, na.rm = TRUE),
    
    median_gross_volumetric_change = median(gross_volumetric_change, na.rm = TRUE),
    mean_gross_volumetric_change = mean(gross_volumetric_change, na.rm = TRUE),
    
    median_abs_net_volumetric_change = median(abs(net_volumetric_change), na.rm = TRUE),
    mean_abs_net_volumetric_change = mean(abs(net_volumetric_change), na.rm = TRUE),
    
    median_directionality = median(directionality, na.rm = TRUE),
    mean_directionality = mean(directionality, na.rm = TRUE),
    
    .groups = "drop"
  ) %>%
  arrange(interface, ksfloor_val)

interface_diagnostics_summary_full

write_csv(interface_diagnostics_summary_full, path(cache_dir, "interface_diagnostics_summary_full.csv"))
```

```{r}
# Summary of annual layer states over the full simulation

layer_diagnostics_summary_full <- df_layers_annual %>%
  group_by(
    scenario,
    ksfloor_factor,
    ksfloor_val,
    variable,
    layer
  ) %>%
  summarise(
    n_years = n(),
    
    median_value = median(mean_value, na.rm = TRUE),
    mean_value_full = mean(mean_value, na.rm = TRUE),
    min_annual_mean = min(mean_value, na.rm = TRUE),
    max_annual_mean = max(mean_value, na.rm = TRUE),
    sd_annual_mean = sd(mean_value, na.rm = TRUE),
    
    .groups = "drop"
  ) %>%
  arrange(variable, layer, ksfloor_val)

layer_diagnostics_summary_full
```

```{r}
layer_diagnostics_summary_full %>%
  filter(variable == "SWC") %>%
  select(
    ksfloor_factor,
    layer,
    median_value,
    min_annual_mean,
    max_annual_mean,
    sd_annual_mean
  ) %>%
  arrange(layer, ksfloor_factor)

df_layers_annual %>%
  filter(
    variable == "SWC",
    ksfloor_factor != "1e_10"
  ) %>%
  ggplot(aes(x = year, y = mean_value, color = ksfloor_factor)) +
  geom_line(linewidth = 0.7) +
  facet_wrap(~ layer, scales = "free_y") +
  labs(
    title = "Annual mean SWC by soil layer",
    subtitle = "Redprec scenario, excluding ksfloor = 1e-10",
    x = "Simulation year",
    y = "Annual mean SWC",
    color = "ksfloor"
  ) +
  theme_bw()

df_layers_annual %>%
  filter(
    variable == "SWP",
    !ksfloor_factor %in% c("1e_10", "1e_12")
  ) %>%
  ggplot(aes(x = year, y = mean_value, color = ksfloor_factor)) +
  geom_line(linewidth = 0.7) +
  facet_wrap(~ layer, scales = "free_y") +
  labs(
    title = "Annual mean SWP by soil layer",
    subtitle = "Redprec scenario, ksfloor <= 1e-14",
    x = "Simulation year",
    y = "Annual mean SWP",
    color = "ksfloor"
  ) +
  theme_bw()
```
Preliminary post-spinup cutoff based on visual inspection (100 years)

```{r}
post_spinup_start <- 101

df_layers_post100 <- df_water_balance_annual %>%
  filter(year >= post_spinup_start)

df_layers_post100 %>% 
filter(
    variable == "SWP",
    !ksfloor_factor %in% c("1e_10", "1e_12")
  ) %>%
  ggplot(aes(x = year, y = mean_value, color = ksfloor_factor)) +
  geom_line(linewidth = 0.7) +
  facet_wrap(~ layer, scales = "free_y") +
  labs(
    title = "Annual mean SWP by soil layer",
    subtitle = "Redprec scenario, ksfloor <= 1e-14",
    x = "Simulation year",
    y = "Annual mean SWP",
    color = "ksfloor"
  ) +
  theme_bw()

df_layers_post100 %>% 
filter(
    variable == "SWC",
    !ksfloor_factor %in% c("1e_10", "1e_12")
  ) %>%
  ggplot(aes(x = year, y = mean_value, color = ksfloor_factor)) +
  geom_line(linewidth = 0.7) +
  facet_wrap(~ layer, scales = "free_y") +
  labs(
    title = "Annual mean SWC by soil layer",
    subtitle = "Redprec scenario, ksfloor <= 1e-14",
    x = "Simulation year",
    y = "Annual mean SWC",
    color = "ksfloor"
  ) +
  theme_bw()

```

Biogeochemical variables

```{r}
df_biogeochemical_annual <- readRDS(
  path(cache_dir, "df_biogeochemical_annual.rds")
)

biog_vars <- names(df_biogeochemical_annual) %>%
  setdiff(c("scenario", "ksfloor_val", "ksfloor_factor", "year"))

biog_vars

plot_biog <- ggplot(df_biogeochemical_annual, aes(x = year, y = mean_value, color = ksfloor_factor)) +
  geom_line(linewidth = 0.7) +
  facet_wrap(~ variable, scales = "free_y") +
  theme_bw() +
  labs(
    x = "Simulation year",
    y = "Annual mean value",
    color = "Ksfloor",
    title = "Annual biogeochemical variables across ksfloor scenarios"
  )

ggsave(
  filename = path(plot_dir, "allbiogvar_fullTrajectory_allKsfloor.png"),
  plot = plot_biog,
  width = 11,
  height = 12,
  dpi = 300
)

```

```{r}
df_biog_plot <- df_biogeochemical_annual %>%
  filter(ksfloor_factor != "1e_10")

biog_variables <- sort(unique(df_biog_plot$variable))

biog_variables

walk(biog_variables, function(var_i) {
  
  p <- df_biog_plot %>%
    filter(variable == var_i) %>%
    ggplot(aes(x = year, y = mean_value, color = ksfloor_factor)) +
    geom_line(linewidth = 0.7) +
    theme_bw() +
    labs(
      title = paste("Annual", var_i, "across ksfloor scenarios"),
      subtitle = "Excluding ksfloor = 1e-10",
      x = "Simulation year",
      y = "Annual mean value",
      color = "Ksfloor"
    )
  
  filename_i <- paste0(
    "biogeochemical_",
    str_replace_all(var_i, "[^A-Za-z0-9_]+", "_"),
    "_fullTrajectory_ksfloor_gt10.png"
  )
  
  ggsave(
    filename = path(plot_dir, filename_i),
    plot = p,
    width = 8,
    height = 5,
    dpi = 300
  )
})
```