---
title: "Bianca"
author: "S. Schmitt"
date: last-modified
format: 
  html:
    embed-resources: true
    toc: true
---

# Setup

You only need the `tidyverse` package for table manipulation and figures preparation. 

```{r libs}
#| message: false
#| warning: false
library(tidyverse)
```

The parameters are the vertical resolution of the soil depth (here 10 cm),
the layer depths used in the model,
and the dry and wet season month used for inter annual summary.

> Note that I used end of the rain season where precipiation is not the highest but soil water is.

```{r pars}
#| message: false
#| warning: false
resolution <- 0.1
layers_depths_shallow <- c(0.1, 0.2, 0.4, 0.7, 1.1, 2.5)
layers_depths_deep <- c(0.1, 0.5, 1.0, 2.4, 6.0, 10.0)
dry_months <- 10:11
wet_months <- 6:7
```

# Data

First we need to compute layers depth structure with cumulated minimum and maximum values.

```{r depths}
#| message: false
#| warning: false
depths_shallow <- tibble(depth = layers_depths_shallow) %>%
  mutate(depth_max = cumsum(depth)) %>% 
  mutate(depth_min = lag(depth_max)) %>% 
  mutate(depth_min = ifelse(is.na(depth_min), 0, depth_min))

depths_deep <- tibble(depth = layers_depths_deep) %>%
  mutate(depth_max = cumsum(depth)) %>% 
  mutate(depth_min = lag(depth_max)) %>% 
  mutate(depth_min = ifelse(is.na(depth_min), 0, depth_min))

```

Then we open the TROLL output of water balance.
For test purposes we reduce here to the years 2005 to 2010.
And we extract the soil water content and potentials (SWC & SWP) and add the corresponding layer depths.

```{r data}
#| message: false
#| warning: false
model_path_shallow <- "~/Desktop/Postdoc_Toulouse/Postdoc_Toulouse/TROLL/TROLL_code_understanding_documenting/runs/sensitivity_tests/shallow_thick_sandy"
model_path_deep <- "~/Desktop/Postdoc_Toulouse/Postdoc_Toulouse/TROLL/TROLL_code_understanding_documenting/runs/sensitivity_tests/deep_thick_sandy"
file <- "(null)_0_water_balance.txt"
file_path_shallow <- file.path(model_path_shallow, file)
file_path_deep <- file.path(model_path_deep, file)

data_shallow <- read_tsv(file_path_shallow) %>% 
  mutate(date = as_date("2004/01/01") + iter) %>% 
  # filter(year(date) %in% 2005:2010) %>% 
  select(date, starts_with("SWC"), starts_with("SWP")) %>% 
  gather(variable, value, -date) %>% 
  separate(variable, c("variable", "layer"), convert = TRUE) %>% 
  mutate(depth_max = depths_shallow$depth_max[layer+1]) %>% 
  mutate(depth_min = depths_shallow$depth_min[layer+1]) %>% 
  select(date, depth_min, depth_max, variable, value) %>% 
  mutate(simulation = "shallow")
data_deep <- read_tsv(file_path_deep) %>% 
  mutate(date = as_date("2004/01/01") + iter) %>% 
  # filter(year(date) %in% 2005:2010) %>% 
  select(date, starts_with("SWC"), starts_with("SWP")) %>% 
  gather(variable, value, -date) %>% 
  separate(variable, c("variable", "layer"), convert = TRUE) %>% 
  mutate(depth_max = depths_deep$depth_max[layer+1]) %>% 
  mutate(depth_min = depths_deep$depth_min[layer+1]) %>% 
  select(date, depth_min, depth_max, variable, value) %>% 
  mutate(simulation = "deep")

data <- bind_rows(data_shallow, data_deep)
```

To compute figures at a finer vertical resolution,
we first need to round up the depth to the resolution we will use,
and then we simply expand observed values to sublayers.

```{r raster}
#| message: false
#| warning: false
raster <- data %>% 
  mutate(depth_min = round(depth_min, 1),
         depth_max = round(depth_max, 1)) %>% 
  rowwise() %>% 
  mutate(depth = list(seq(depth_min, 
                          depth_max, 
                          by = resolution))) %>% 
  unnest(depth)
```

# Figures

And here are the codes to represent SWC and SWP as raster with depth and time or
mean dry and wet season profiles.

## Rasters

```{r swc_rast}
#| message: false
#| warning: false
#| fig-cap: "Soil water content across time (x-axis) and depth (y-axis)."
raster %>% 
  filter(variable == "SWC") %>% 
  ggplot(aes(date, depth, fill = value)) +
  geom_raster() +
  theme_bw() +
  scale_y_reverse() +
  scale_fill_viridis_c(expression("SWC ["~m^3~m^3~"]"),
                       direction = -1) +
  xlab("") +
  ylab("Depth [ m ]")+
  facet_wrap(~simulation, nrow = 1)
```

```{r swp_rast}
#| message: false
#| warning: false
#| fig-cap: "Soil water potential across time (x-axis) and depth (y-axis)."
raster %>% 
  filter(variable == "SWP") %>% 
  ggplot(aes(date, depth, fill = -value)) +
  geom_raster() +
  theme_bw() +
  scale_y_reverse() +
  scale_fill_viridis_c(expression("|SWP| ["~MPa~"]"),
                       direction = 1, trans = "log") +
  xlab("") +
  ylab("Depth [ m ]")
```

## Profiles

> Profiles are made with daily means across years, but maybe we should use first a season mean within year and then summarise it across years.

```{r swc_profile}
#| message: false
#| warning: false
#| fig-cap: "Mean soil water content for wet and dry season across depth (y-axis). Point represent the median across days and bars the 90% interval."
data %>% 
  filter(variable == "SWC") %>% 
  filter(month(date) %in% c(dry_months, wet_months)) %>% 
  mutate(season = ifelse(month(date) %in% dry_months, "dry", "wet")) %>% 
  group_by(simulation,variable, depth_min, depth_max, season) %>% 
  summarise(m = median(value), 
            ll = quantile(value, .05), 
            hh = quantile(value, .95)) %>% 
  mutate(depth = (depth_min + depth_max)/2) %>% 
  ggplot(aes(depth_max, m, col = paste(simulation, season))) +
  geom_line() +
  geom_linerange(aes(ymin = ll, ymax = hh)) +
  geom_point() +
  theme_bw() +
  coord_flip() +
  scale_x_reverse("Depth [ m ]") +
  ylab(expression("SWC ["~m^3~m^3~"]"))
```

```{r swp_profile}
#| message: false
#| warning: false
#| fig-cap: "Mean soil water potential for wet and dry season across depth (y-axis). Point represent the median across days and bars the 90% interval."
data %>% 
  filter(variable == "SWP") %>% 
  filter(month(date) %in% c(dry_months, wet_months)) %>% 
  mutate(season = ifelse(month(date) %in% dry_months, "dry", "wet")) %>% 
  group_by(simulation,variable, depth_min, depth_max, season) %>% 
  mutate(value = -value) %>% 
  summarise(m = median(value), 
            ll = quantile(value, .05), 
            hh = quantile(value, .95)) %>% 
  mutate(depth = (depth_min + depth_max)/2) %>% 
  ggplot(aes(depth, m, col = paste(simulation,season))) +
  geom_line() +
  geom_linerange(aes(ymin = ll, ymax = hh)) +
  geom_point() +
  theme_bw() +
  coord_flip() +
  scale_x_reverse("Depth [ m ]") +
  scale_y_log10(expression("|SWP| ["~MPa~"]"))
```

