

```{r}
knitr::opts_chunk$set(echo = TRUE)
# Loading required packages
library(tidyverse)
library(data.table) 
library(fs)         
```

```{r}
# Load data
# script to generate the .rds files:
# "~/Desktop/Postdoc_Toulouse/Postdoc_Toulouse/TROLL/TROLL_code_understanding_documenting/runs/reserva_Ducke/save_data_rds.R"

project_dir <- path_expand(
  "~/Desktop/Postdoc_Toulouse/Postdoc_Toulouse/TROLL/TROLL_code_understanding_documenting/runs/reserva_Ducke"
)

cache_dir <- path(project_dir, "_rds")

plot_dir <- path(project_dir, "plots")
dir_create(plot_dir)

# Daily data
# df_water_balance_daily <- readRDS(path(cache_dir, "df_water_balance_daily.rds"))
# df_vertical_flux_daily <- readRDS(path(cache_dir, "df_vertical_flux_daily.rds"))
# df_biogem_daily <- readRDS(path(cache_dir, "df_biogem_daily.rds"))

# Annual data
df_water_balance_annual <- readRDS(path(cache_dir, "df_water_balance_annual.rds"))
df_vertical_flux_annual <- readRDS(path(cache_dir, "df_vertical_flux_annual.rds"))
df_biogem_annual <- readRDS(path(cache_dir, "df_biogeochemical_annual.rds"))
```

Full Timeline and Stabilization Point

Before looking at the final state, we need to determine *if* and *when* the system stabilizes. A hydrological model is considered stable (having surpassed the spin-up phase) 

# The Full 274-Year Trajectory

First, let's plot the entire time series to visually identify the spin-up period and whether the curves flatten out.

To consider in the analysis:
Increasing ksfloor reduces the risk of numerical hydraulic locking by preventing extremely low hydraulic conductivity values. However, higher ksfloor values also impose an increasingly artificial lower bound on soil-water movement. If the imposed floor is too high, vertical connectivity may be maintained even when the soil profile should physically become hydraulically disconnected, especially under deep water table and sandy soil conditions. Therefore, the sensitivity analysis should identify the lowest ksfloor that stabilizes the model without artificially increasing vertical water exchange.

```{r plot-full-swc-trajectory}
swc_full <- df_water_balance_annual %>%
  filter(variable == "SWC") %>%
  ggplot(aes(x = year, y = mean_value, color = scenario)) +
  geom_line(linewidth = 0.7) +
  facet_wrap(~ layer, scales = "free_y") +
  labs(
    title = "SWC",
    subtitle = "Reserva Ducke scenarios",
    x = "Simulation year",
    y = "Annual mean SWC",
    color = "Scenario"
  ) +
  theme_bw()

swc_full

ggsave(
  filename = path(plot_dir, "SWC_fullTrajectory_Ducke_scenarios.png"),
  plot = swc_full,
  width = 10,
  height = 7,
  dpi = 300
)

```

```{r plot-full-swp-trajectory}

swp_full_allks <-  df_water_balance_annual %>%
  filter(variable == "SWP") %>%
  ggplot(aes(x = year, y = mean_value, color = scenario)) +
  geom_line(linewidth = 0.7) +
  facet_wrap(~ layer, scales = "free_y") +
  labs(
    title = "SWP",
    x = "Simulation year",
    y = "Annual mean SWP",
    color = "Scenario"
  ) +
  theme_bw()

swp_full_allks

ggsave(
  filename = path(plot_dir, "SWP_fullTrajectory_allKsfloor.png"),
  plot = swp_full_allks,
  width = 10,
  height = 7,
  dpi = 300
)

```

```{r plot-full-transpiration-trajectory}

transpiration_full_allks <- df_water_balance_annual %>%
  filter(variable == "transpiration") %>%
  ggplot(aes(x = year, y = mean_value, color = scenario)) +
  geom_line(linewidth = 0.7) +
  facet_wrap(~ layer, scales = "free_y") +
  labs(
    title = "Transpiration",
    x = "Simulation year",
    y = "Annual mean transpiration",
    color = "Scenario"
  ) +
  theme_bw()

transpiration_full_allks

ggsave(
  filename = path(plot_dir, "transpiration_fullTrajectory_allKsfloor.png"),
  plot = transpiration_full_allks,
  width = 10,
  height = 7,
  dpi = 300
)

```



Plotting layer interface variables

```{r}
mean_ksgeometric_full_allks <- df_vertical_flux_annual %>%
  filter(metric == "mean_ks_harmonic") %>% # actually this is the ks geometric
  ggplot(aes(x = year, y = mean_value, color = scenario)) +
  geom_line(linewidth = 0.7) +
  facet_wrap(~ interface, scales = "free_y") +
  scale_y_log10() +
  labs(
    title = "Mean harmonic hydraulic conductivity by interface",
    subtitle = "Annual mean values",
    x = "Simulation year",
    y = "Annual mean Ks geometric",
    color = "Scenario"
  ) +
  theme_bw()

mean_ksgeometric_full_allks

ggsave(
  filename = path(plot_dir, "mean_ksgeometric_fullTrajectory_allKsfloor.png"),
  plot = mean_ksgeometric_full_allks,
  width = 10,
  height = 7,
  dpi = 300
)

```


```{r}
gross_volumetricchange_full_allks <- df_vertical_flux_annual %>%
  filter(metric == "gross_volumetric_change") %>%
  ggplot(aes(x = year, y = sum_value, color = scenario)) +
  geom_line(linewidth = 0.7) +
  facet_wrap(~ interface, scales = "free_y") +
  labs(
    title = "Gross vertical water exchange by interface",
    subtitle = "Annual sums",
    x = "Simulation year",
    y = "Annual gross volumetric change",
    color = "Scenario"
  ) +
  theme_bw()

gross_volumetricchange_full_allks

ggsave(
  filename = path(plot_dir, "gross_volumetricchange_fullTrajectory_allKsfloor.png"),
  plot = gross_volumetricchange_full_allks,
  width = 10,
  height = 7,
  dpi = 300
)

```


```{r}
# Log scale
gross_volumetricchange_full_ksgt10_log <- df_vertical_flux_annual %>%
  mutate(
    sum_value_plot = if_else(sum_value <= 0, NA_real_, sum_value)
  ) %>%
  ggplot(aes(x = year, y = sum_value_plot, color = scenario)) +
  geom_line(linewidth = 0.7, na.rm = TRUE) +
  facet_wrap(~ interface, scales = "free_y") +
  scale_y_log10() +
  labs(
    title = "Gross vertical water exchange by interface",
    subtitle = "Annual sums, log scale",
    x = "Simulation year",
    y = "Annual gross volumetric change, log scale",
    color = "Scenario"
  ) +
  theme_bw()

gross_volumetricchange_full_ksgt10_log

ggsave(
  filename = path(plot_dir, "gross_volumetricchange_fullTrajectory_Ksfloor_gt10_log.png"),
  plot = gross_volumetricchange_full_ksgt10_log,
  width = 10,
  height = 7,
  dpi = 300
)

```


```{r}
net_volumetricchange_full_ksgt10 <- df_vertical_flux_annual %>%
  filter(
    metric == "net_volumetric_change"
  ) %>%
  ggplot(aes(x = year, y = sum_value, color = scenario)) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_line(linewidth = 0.7) +
  facet_wrap(~ interface, scales = "free_y") +
  labs(
    title = "Net vertical water exchange by interface",
    subtitle = "Annual sums",
    x = "Simulation year",
    y = "Annual net volumetric change",
    color = "Scenario"
  ) +
  theme_bw()

net_volumetricchange_full_ksgt10

ggsave(
  filename = path(plot_dir, "net_volumetricchange_fullTrajectory_Ksfloor_gt10.png"),
  plot = net_volumetricchange_full_ksgt10,
  width = 10,
  height = 7,
  dpi = 300
)


```

Looking to directionality
```{r}
df_directionality_annual <- df_vertical_flux_annual %>%
  filter(metric %in% c("net_volumetric_change", "gross_volumetric_change")) %>%
  select(
    scenario,
    year,
    interface,
    layer_upper,
    layer_lower,
    metric,
    sum_value
  ) %>%
  pivot_wider(
    names_from = metric,
    values_from = sum_value
  ) %>%
  mutate(
    abs_net_volumetric_change = abs(net_volumetric_change),
    directionality = abs_net_volumetric_change / gross_volumetric_change,
    directionality = if_else(
      is.finite(directionality),
      directionality,
      NA_real_
    )
  )

volume_directionality_full_ksgt10 <- df_directionality_annual %>%
  ggplot(aes(x = year, y = directionality, color = scenario)) +
  geom_line(linewidth = 0.7) +
  facet_wrap(~ interface, scales = "free_y") +
  labs(
    title = "Directionality of vertical water exchange by interface",
    subtitle = "|net volumetric change| / gross volumetric change",
    x = "Simulation year",
    y = "Directionality",
    color = "Scenario"
  ) +
  theme_bw()

volume_directionality_full_ksgt10

ggsave(
  filename = path(plot_dir, "volume_directionality_fullTrajectory_Ksfloor_gt10.png"),
  plot = volume_directionality_full_ksgt10,
  width = 10,
  height = 7,
  dpi = 300
)

# directionality ≈ 1
# fluxo anual quase totalmente em uma direção.
# 
# directionality ≈ 0
# há troca nos dois sentidos que se cancela, ou net muito pequeno em relação ao gross.
# 
# directionality = NA
# gross volumetric change é zero ou quase zero, então não há troca suficiente para calcular.

```



Post spinup
```{r}
post_spinup_start <- 101

df_layers_post100 <- df_water_balance_annual %>%
  filter(year >= post_spinup_start)

df_layers_post100 %>% 
filter(
    variable == "SWP",
    !ksfloor_factor %in% c("1e_10", "1e_12")
  ) %>%
  ggplot(aes(x = year, y = mean_value, color = ksfloor_factor)) +
  geom_line(linewidth = 0.7) +
  facet_wrap(~ layer, scales = "free_y") +
  labs(
    title = "Annual mean SWP by soil layer",
    subtitle = "Redprec scenario, ksfloor <= 1e-14",
    x = "Simulation year",
    y = "Annual mean SWP",
    color = "ksfloor"
  ) +
  theme_bw()

df_layers_post100 %>% 
filter(
    variable == "SWC",
    !ksfloor_factor %in% c("1e_10", "1e_12")
  ) %>%
  ggplot(aes(x = year, y = mean_value, color = ksfloor_factor)) +
  geom_line(linewidth = 0.7) +
  facet_wrap(~ layer, scales = "free_y") +
  labs(
    title = "Annual mean SWC by soil layer",
    subtitle = "Redprec scenario, ksfloor <= 1e-14",
    x = "Simulation year",
    y = "Annual mean SWC",
    color = "ksfloor"
  ) +
  theme_bw()

```


Biogeochemical variables

```{r}
df_biogeochemical_annual <- readRDS(
  path(cache_dir, "df_biogeochemical_annual.rds")
)

biog_vars <- names(df_biogeochemical_annual) %>%
  setdiff(c("scenario", "year"))

biog_vars

plot_biog <- ggplot(
  df_biogeochemical_annual,
  aes(x = year, y = mean_value, color = scenario, group = scenario)
) +
  geom_ribbon(
    aes(
      ymin = mean_value - sd_value,
      ymax = mean_value + sd_value,
      fill = scenario
    ),
    alpha = 0.25,
    color = NA
  ) +
  geom_line(linewidth = 0.3) +
  facet_wrap(~ variable, scales = "free_y") +
  theme_bw() +
  labs(
    x = "Simulation year",
    y = "Annual mean value",
    color = "Scenario",
    fill = "Scenario",
    title = "Annual biogeochemical variables",
    subtitle = "Shaded area = annual mean ± SD"
  )

plot_biog

ggsave(
  filename = path(plot_dir, "allbiogvar_fullTrajectory_withSD.png"),
  plot = plot_biog,
  width = 7,
  height = 12,
  dpi = 300
)
```

```{r}
df_biog_plot <- df_biogeochemical_annual

biog_variables <- sort(unique(df_biog_plot$variable))

biog_variables

walk(biog_variables, function(var_i) {
  
  p <- df_biog_plot %>%
    filter(variable == var_i) %>%
    ggplot(aes(x = year, y = mean_value, color = scenario)) +
    geom_line(linewidth = 0.7) +
    theme_bw() +
    labs(
      title = paste("Annual", var_i),
      subtitle = "Reserva Ducke scenarios",
      x = "Simulation year",
      y = "Annual mean value",
      color = "Scenario"
    )
  
  filename_i <- paste0(
    "biogeochemical_",
    str_replace_all(var_i, "[^A-Za-z0-9_]+", "_"),
    "_fullTrajectory_Ducke_scenarios.png"
  )
  
  ggsave(
    filename = path(plot_dir, filename_i),
    plot = p,
    width = 8,
    height = 5,
    dpi = 300
  )
})
```

